---
title: "Trees and forests"
output: github_document
---

# import libraries

```{r}
library(tidyverse)
library(caret)
library(pROC)
```

Build custom AUC function to extract AUC from the caret model object

```{r}
eval_mod <- function(model, data) {
  pred <- predict(model, data)
  cm <- caret::confusionMatrix(pred, data$classes, positive="malignant")
  auc <- roc(data$classes,
             predict(model, data, type = "prob")[, "malignant"]) %>% auc()
  result <- c(cm$overall["Accuracy"],cm$byClass['Sensitivity'], cm$byClass['Specificity'], cm$byClass['F1'],AUC=auc)
  return(result)
}
```


# Read clean data 

```{r}
bc_data <- readRDS("../EDA/bc_clean.RDS")
bc_data$classes <- as.factor(bc_data$classes)
```

```{r}
set.seed(2024)
index <- caret::createDataPartition(bc_data$classes, p = 0.7, list = FALSE)

train_data <- bc_data[index, ]
test_data  <- bc_data[-index, ]
```

```{r}
train_data$classes %>% table(.)
set.seed(2024)
library(rpart)
tree_mod <- rpart(
  formula = classes ~. ,
  data    = train_data,
  method  = "class"
)
tree_mod
```

```{r}
library(rpart.plot)
rpart.plot(tree_mod)
```

# Pruning
We prune the tree using the cost complexity criterion. Can a less deep tree give comparable results. If so, go with the shallower tree to reduce the likelihood of overfitting. We used the a parameter called complexity parameter (CP)

```{r}
printcp(tree_mod)
```

It computes the cross validation error for each value of cp CP=0.01 gives the lowest error

```{r}
opt <- which.min(tree_mod$cptable[,"xerror"])
```

get its value

```{r}
cp <- tree_mod$cptable[opt, "CP"]
cp
```

We can prune the tree based on this CP
```{r}
pruned_model <- prune(tree_mod,cp)
#plot tree
rpart.plot(pruned_model)
```
#Note that rpart will use a default CP value of 0.01 if you don’t specify one in prune.

```{r}

pred <- predict(object = tree_mod,   # model object 
                newdata = test_data,
                type="class")  # test dataset
pruned_tree_metrics <- caret::confusionMatrix(pred,test_data$classes, positive = "malignant")
pruned_tree_metrics
```

```{r}
set.seed(2024)
caret_tree <- train(
  classes ~ .,
  data = train_data,
  method = "rpart",
  metric ="ROC",
  trControl = trainControl(method = "cv", number = 5,
                           classProbs = T, summaryFunction = twoClassSummary),
  tuneLength = 20
)
ggplot(caret_tree)
```
```{r}
caret_tree$bestTune
rpart.plot(caret_tree$finalModel)
```
```{r}
tree <- eval_mod(caret_tree,test_data)
tree
```
```{r}
tree_bag1 <- ipred::bagging(
formula = classes ~ .,
data = train_data,
nbagg = 500,  
coob = TRUE
)
tree_bag1
```
```{r}
set.seed(2024)
caret_bag <- train(
  classes ~ .,
  data = train_data,
  method = "treebag",
  trControl = trainControl(method = "cv", number = 5,
                           classProbs = T, summaryFunction = twoClassSummary),
  metric ="ROC",
  nbagg = 20,  
  control = rpart.control(minsplit = 2, cp = 0)
)
caret_bag

bag <- eval_mod(caret_bag,test_data)
rbind(tree, bag)
```

# Random Forest

```{r}
library(randomForest)
# Train a Random Forest
set.seed(2024)  # for reproducibility
rf_model <- randomForest(formula = classes ~ ., 
                         data = train_data)

# Print the model output                             
print(rf_model)
```
```{r}
caret_rf <- train(
  classes ~ .,
  data = train_data,                         
  method = "ranger",
  metric = "ROC",
  trControl = trainControl(method = "cv", number = 5, classProbs = TRUE, summaryFunction = twoClassSummary),
  importance="impurity"
)

forest <- eval_mod(caret_rf,test_data)

roc_tree <- roc(test_data$classes,
             predict(pruned_model, test_data, type = "prob")[, "malignant"]) %>% auc()
pruned_tree <- c(pruned_tree_metrics$overall['Accuracy'], pruned_tree_metrics$byClass['Sensitivity'],
                 pruned_tree_metrics$byClass['Specificity'], pruned_tree_metrics$byClass['F1'],  "AUC" = roc_tree)
rbind(pruned_tree, tree, bag, forest)
```
