---
title: "Logistic regression"
output: github_document
---

# Importing libraries

```{r}
library(tidyverse) # for tidy data analysis
library(readr)     # for fast reading of input files
library(caret)
library(pROC)
```

# Function for extracting metrics 
```{r}
eval_mod <- function(model, data) {
  pred <- predict(model, data)
  cm <- caret::confusionMatrix(pred, data$classes, positive="malignant")
  auc <- roc(data$classes,
             predict(model, data, type = "prob")[, "malignant"]) %>% auc()
  result <- c(cm$overall["Accuracy"],cm$byClass['Sensitivity'], cm$byClass['Specificity'], cm$byClass['F1'],AUC=auc)
  return(result)
}
```

# Importing data

```{r}
bc_data0 <-  read.csv(paste0("http://archive.ics.uci.edu/ml/machine-learning-databases/","breast-cancer-wisconsin/breast-cancer-wisconsin.data"), header = FALSE, stringsAsFactors = F)

names (bc_data0) <-  c("sample_code_number", 
                       "clump_thickness", 
                       "uniformity_of_cell_size", 
                       "uniformity_of_cell_shape", 
                       "marginal_adhesion", 
                       "single_epithelial_cell_size", 
                       "bare_nuclei", 
                       "bland_chromatin", 
                       "normal_nucleoli", 
                       "mitosis", 
                       "classes")
```

# Data wrangling 
```{r}
str(bc_data0)
```

```{r}
bc_data0$bare_nuclei = as.integer(bc_data0$bare_nuclei)
bc_data1 <- bc_data0 %>%
  dplyr::mutate(classes = ifelse(classes == "2", "benign",
                                 ifelse(classes == "4", "malignant", NA)))
```

# De-duplicate observations
```{r}
bc_data2 <- bc_data1 %>% distinct(sample_code_number,.keep_all = TRUE)
row.names(bc_data2) <- bc_data2$sample_code_number
bc_data3 <- bc_data2 %>% select(-sample_code_number)
```

# Split data into training and testing
```{r}
set.seed(123)
train.index <- caret::createDataPartition(bc_data3$classes, p = .7, list=FALSE)

train <- bc_data3[ train.index,]
valid  <- bc_data3[-train.index,]
```

# Check the missing values in the data
```{r}
summary(train)
sapply(train, function(x) sum(is.na(x)))
dim (train)
```

#	Pairwise deletion: ignore missing values

```{r}
mean(train$bare_nuclei)
mean(train$bare_nuclei, na.rm=TRUE)
mean(train$clump_thickness, na.rm=TRUE)
cor(train$bare_nuclei,train$clump_thickness)
cor(train$bare_nuclei,train$clump_thickness, use = "pairwise.complete.obs")
```

# Listwise deletion: Remove  the rows where missing values occur. 
```{r}
listwise <- train[complete.cases(train),]
sapply(listwise, function(x) sum(is.na(x)))
dim(listwise)
mean(listwise$bare_nuclei)
mean(listwise$clump_thickness)
cor(listwise$bare_nuclei,listwise$clump_thickness)
```

# Impute missing value with mean

Multiple imputation by chained equations (MICE): blue is observed, red is imputed
```{r}
Mean_imp <- mice::mice(train, m=5, method = 'mean', print = FALSE)
mice::densityplot(Mean_imp)
# MULTIPLE IMPUTATION USING PREDICTIVE MEAN MATCHING
pmm_imp <- mice::mice(train,m=10, method = "pmm", print = FALSE)
mice::densityplot(pmm_imp)
```
```{r}
imputed<-complete(data=pmm_imp)
mean(imputed$bare_nuclei)
mean(listwise$bare_nuclei)
mean(imputed$clump_thickness)
mean(listwise$clump_thickness)
mean(train$clump_thickness)
```

Will not impute validation
```{r}
sapply(valid, function(x) sum(is.na(x)))
completeValid <- valid[complete.cases(valid),]
table(completeValid$classes)
imputed$classes <- as.factor(imputed$classes)
completeValid$classes <- as.factor(completeValid$classes)
```
# Class Imbalance
```{r}
table(imputed$classes)
```

```{r}
set.seed(123)
ctrl = trainControl(method = "repeatedcv", 
                                number = 5, 
                                repeats = 3, 
                                summaryFunction = twoClassSummary,
                                classProbs = TRUE,
                                verboseIter = FALSE)

orig_fit <- caret::train(classes ~ .,
                  data = imputed,
                  method = "glm",
                  family = "binomial",
                  preProcess = c("scale", "center"),
                  metric ="ROC",
                  trControl = ctrl)
original <- eval_mod(orig_fit, completeValid)
```

# Downsampling
```{r}
set.seed(123)
ctrl$sampling <- "down"
down_fit <- train(classes ~ .,
                  data = imputed,
                  method = "glm",
                  family = "binomial",
                  preProcess = c("scale", "center"),
                  metric = "ROC",
                  trControl = ctrl)
down <- eval_mod(down_fit, completeValid)
```

# Upsampling
```{r}
set.seed(123)
ctrl$sampling <- "up"
up_fit <- train(classes ~ .,
                  data = imputed,
                  method = "glm",
                  family = "binomial",
                  preProcess = c("scale", "center"),
                  metric = "ROC",
                  trControl = ctrl)
up <- eval_mod(up_fit, completeValid)
```

# SMOTE
```{r}
set.seed(123)

ctrl$sampling <- "smote"

smote_fit <- train(classes ~ .,
                   data = imputed,
                   method = "glm",
                   family = "binomial",
                   preProcess = c("scale", "center"),
                   metric = "ROC",
                   trControl = ctrl)
smote <- eval_mod(smote_fit, completeValid)
```

```{r}
Final <- rbind(original, down, up, smote)
Final
```
